---
title: "Grey Relational Analysis(GRA)"
author: "Nisus Liu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Grey Relational Analysis(GRA)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Grey Relational Analysis include two important function.
First function: grey relational degree, which is similar to orrelation coefficient, if you want to evaluate some units, please transpose data frame before using this function. Second funtion: grey clustering, like hierarchical clustering, see `hclust`.

## Grey Relational Degree

There are two usage of grey relational degree. This algorithm is to measure similarity of two variables, just like `cor`. You can transpose your data set if you want to evaluate some units.

* One is to inspect correlation of two variables, data type like:

| reference | v1 | v2 | v3 |
|-----------|----|----|----|
|    1.2    | 1.8|0.9 | 8.4|
|    0.11   | 0.3|0.5 | 0.2|
|    1.3    | 0.7|0.12|0.98|
|    1.9    |1.09|2.8 |0.99|

`reference`: reference variable, grey relational degree between `reference` and `v1`... approximately measures the similarity of `reference` and `v1`.


* Another is to evaluate good or bad about some units.

| units | v1 | v2 | v3 |
|-----------|----|----|----|
|    jiangsu   | 1.8|0.9 | 8.4|
|    zhejiang   | 0.3|0.5 | 0.2|
|    anhui    | 0.7|0.12|0.98|
|    fujian    |1.09|2.8 |0.99|


### Example

```{r, include=FALSE}

#' @title Grey Relational Analysis(GRA)
#' @description first function: grey relational degree, which is similar to orrelation coefficient, if you want to evaluate some unit, please transpose data frame before using this function.
#' second funtion: grey clustering, like hierarchical clustering, see \code{hclust}.
#' @param df a data frame with standardized data which can aviod the effects of dimensions. Note: it must
#' includes reference sequence.
#' @param referColNum the number of reference sequence in \code{df}.
#' @param distingCoeff distinguishing coefficient, the smaller the higher distinguish.
#' @param cluster logical value, indicates whether to cluster, default is \code{FALSE}. Note: this algorithm clusters by column(most clustering method are by row) because it is
#' base on grey relational degree algorithm(mode like correlation coefficient, not evaluating mode).
#' @param clusterMethod method of cluster.

#' @examples

#' ## generate data
#' refer = c(1,1,1,1)
#' liaoning = c(0.064, 0.082,0.978,0.423)
#' shandong = c(0.101,0.3,1,0.917)
#' jiangsu = c(0.114,0.14,0.943, 0.315)
#' zhejiang = c(0.102,0.147,0.934,0.395)
#' fujian = c(0.022,0.053,0.927,0.061)
#' guangdong = c(1,1,0.012,1)
#' economyCompare = data.frame(refer, liaoning, shandong, jiangsu, zhejiang, fujian, guangdong)
#' rownames(economyCompare) = c("indGV", "indVA", "profit", "incomeTax")

#' ## Grey Relational Degree
#' greyRelDegree = GRA(economyCompare)
#' greyRelDegree

#' ## Grey Clustering
#' GRA(economyCompare, cluster = T)


GRA = function(df, referColNum = 1, distingCoeff = 0.5, cluster = FALSE, clusterMethod = "single")  {
  
  # 异常控制 #
  if (any(is.na(df))) stop("'df' have NA" )
  if (distingCoeff<0 | distingCoeff>1) stop("'distingCoeff' must be in range of [0,1]" )
  if (referColNum<0 | referColNum>ncol(df)) stop("'referColNum' out of range")
  
  
  y = df[referColNum]
  nr = nrow(y)
  X = df[-referColNum]
  nc = ncol(X)
  
  
  diff = X  #设置差学列矩阵空间
  
  for (i in 1:nc) {
    diff[i] = abs(y-X[i])
  }
  
  mi = min(diff)
  mx = max(diff)
  
  
  #计算关联系数#
  relations = (mi+distingCoeff*mx) / (diff + distingCoeff*mx)
  
  #计算关联度#
  # 暂时简单处理, 等权
  relDegree = rep(NA, nc)
  for (i in 1:nc) {
    relDegree[i] = mean(relations[,i])  # 等权
  }
  
  
  #排序: 按关联度大到小#
  X_order = X[order(relDegree, decreasing = TRUE)]
  relationalDegree = relDegree[order(relDegree, decreasing = TRUE)]
  relDes = rep(NA, nc) #分配空间  关联关系描述(说明谁和谁的关联度)
  X_names = names(X_order)
  for (i in 1:nc) {
    relDes[i] = paste(names(df)[referColNum], X_names[i], sep = "~")
  }
  names(relationalDegree) = relDes
  
  # greyRelDegree = list(relationalDegree = relationalDegree, X_order = X_order)
  
  
  if (cluster) {
    
    greyRelDegree = GRA(economyCompare, referColNum = 1, distingCoeff = 0.5)
    
    
    # relationalDegree = greyRelDegree$relationalDegree
    # X = X_order
    
    
    
    nc = length(relationalDegree)
    
    # 得到差异率矩阵 #
    grey_diff = matrix(0, nrow = nc, ncol = nc)
    for (i in 1:nc) {
      for(j in 1:nc)  {
        
        grey_diff[i,j] = abs(relationalDegree[i] - relationalDegree[j]) / relationalDegree[j]
      }
    }
    #得到距离矩阵#
    grey_dist = matrix(0, nrow = nc, ncol = nc)
    for (i in 1:nc) {
      for (j in 1:nc) {
        grey_dist[i,j] = grey_diff[i,j]+grey_diff[j,i]
      }
    }
    
    # 得到灰色相关系数矩阵 #
    grey_dist_max = max(grey_dist)
    grey_correl = matrix(0, nrow = nc, ncol = nc)
    for (i in 1:nc) {
      for (j in 1:nc) {
        grey_correl[i,j] = 1 - grey_dist[i,j] / grey_dist_max
      }
    }
    
    rownames(grey_correl) = names(X_order)
    colnames(grey_correl) = names(X_order)
    grey_correl
    
    
    d = as.dist(1-grey_correl)  # 得到无对角线的下三角矩阵(数值意义反向了, 值越小表示越相关 )
    # 主对角线其实表示了各个对象的相近程度, 画图的时候, 相近的对象放在一起
    
    hc = hclust(d, method = clusterMethod)  # 系统聚类(分层聚类)函数, single: 单一连接(最短距离法/最近邻) 
    # hc$height, 是上面矩阵的对角元素升序
    # hc$order, 层次树图上横轴个体序号
    plot(hc,hang=-1)  #hang: 设置标签悬挂位置
    
  }
  
  #output#
  
  if (cluster)  {
    lst = list(relationalDegree=relationalDegree, greyCorrelCoeff = grey_correl, X = X_order, referColumn = y)
    
  }
  else {
    lst = list(relationalDegree=relationalDegree, X = X_order, referColumn = y)
  }
  
  return(lst)
  
}


```



```{r}
## generate data
refer = c(1,1,1,1)
liaoning = c(0.064, 0.082,0.978,0.423)
shandong = c(0.101,0.3,1,0.917)
jiangsu = c(0.114,0.14,0.943, 0.315)
zhejiang = c(0.102,0.147,0.934,0.395)
fujian = c(0.022,0.053,0.927,0.061)
guangdong = c(1,1,0.012,1)
economyCompare = data.frame(refer, liaoning, shandong, jiangsu, zhejiang, fujian, guangdong)
rownames(economyCompare) = c("indGV", "indVA", "profit", "incomeTax")

## Grey Relational Degree
greyRelDegree = GRA(economyCompare)
greyRelDegree


```



## Grey Clustering


### Example

```{r}
## Grey Clustering
GRA(economyCompare, cluster = T)
```


